How is experience transformed into memory?

### 2018-08-28 paper/main.tex
supp materials edits

\begin{suppfigure}[t!]
\centering
\includegraphics[width=1\textwidth]{figs/supp4_feature_value.pdf}
\caption{\small \textbf{Impact of individual features on topic modeling analysis.} Contribution of each feature to model structure. Bars represent the correlation of a video model trained in the absence of a given feature to the model trained on all features. B).Contribution of each feature to video model/recall model relationship. The leftmost bar represents the across-subjects mean correlation between the video model and recall models trained on all features. Subsequent bars represent the same relationship between video and recall models trained in the absence of a given feature. Error bars are the standard error of the mean across subjects.}
\label{fig:feature-importance}
\end{suppfigure}

\section*{Naturalistic extensions of classic list-learning analyses}
Just like in a traditional ``free-recall'' list-learning experiment where participants view a list of words and then verbally recall them, our video-recall matching analysis approach affords us the ability to analyze memory in the same way. The recalled events can be treated as ``items'' analogous to words recalled in a list-learning study. Here, we sought to characterize memory performance/dynamics by extending classic analyses originally designed for list-learning experiments to more naturalistic settings.

@@

\section*{Additional measures of naturalistic memory}
To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the video model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered and computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events our of order or with low precision the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants ($p<.05$, participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to hand annotated memory performance across participants (Pearson's $r(16)=-.57, p=.016$). Thus, a closer match between the video and recall event models was related to better recall performance.

@@

\section*{Supplemental Methods}
\subsection{Quantifying the importance of features}
To determine the contribution of each feature to the structure of the video model, we examined the similarity between the temporal structure of models trained in the absence of a single feature and that of our original (i.e. full) model. First, we iteratively removed one transcribed feature from the scene descriptions and constructed a timepoints (1976) by topis (100) matrix using a topic model fit to the remaining features. We then represented the original model as well as the new model’s temporal structure as a timepoints (1976) by timepoints (1976) correlation matrix. Finally, we vectorized these correlation matrices and correlated them to each other resulting in a single number (for each feature removed from the model) representing the similarity of the ``feature-removed'' models to the full model.

In order to ascertain which features were important in relating the recall models to the video model, we similarly compared to the temporal structures of video and recall models deprived of a single feature at a time. For each feature removed, we transformed each subject’s recall transcript using a model trained on the feature-deprived video text windows (of 50 scene segments), and resampled the recall timeseries to match the shape of the video model (1976 timepoints). We then represented the temporal structures of each subject’s recall model as timepoints (1976) by timepoints (1976) correlation matrices and computed the average correlation with the temporal structure of the video model (across subjects).

>>>>

\begin{suppfigure}[t!]
\centering
\includegraphics[width=1\textwidth]{figs/supp4_feature_value.pdf}
\caption{\small \textbf{Impact of individual features on topic modeling analysis.} \textbf{A.} Contribution of each feature to model structure. Bars represent the correlation of a video model trained in the absence of a given feature to the model trained on all features. \textbf{B.} Contribution of each feature to video model/recall model relationship. The leftmost bar represents the across-subjects mean correlation between the video model and recall models trained on all features. Subsequent bars represent the same relationship between video and recall models trained in the absence of a given feature. Error bars are the standard error of the mean across subjects.}
\label{fig:feature-importance}
\end{suppfigure}

\section*{Naturalistic extensions of classic list-learning analyses}
Just like in a traditional free recall list-learning experiment where participants view a list of words and then verbally recall them, our video-recall matching analysis approach affords us the ability to analyze memory in the same way. The recalled events can be treated as ``items'' analogous to words recalled in a list-learning study. Here, we sought to characterize memory performance/dynamics by extending classic analyses originally designed for list-learning experiments to more naturalistic settings.

@@

\section*{Additional measures of naturalistic memory}
To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the video model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered and computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events out of order, or with low precision, the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants ($p<.05$, participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to hand annotated memory performance across participants (Pearson's $r(16)=-.57, p=.016$). Thus, a closer match between the video and recall event models was related to better recall performance.

@@

\section*{Supplemental Methods}
\subsection{Quantifying the importance of features}
To determine the contribution of each feature to the structure of the video model, we examined the similarity between the temporal structure of models trained in the absence of a single feature and that of our original (i.e. full) model (Supp. Fig.~\ref{fig:feature-importance}A). First, we iteratively removed one transcribed feature from the scene descriptions and constructed a timepoints (1976) by topis (100) matrix using a topic model fit to the remaining features. We then represented the original model as well as the new model’s temporal structure as a timepoints-by-timepoints correlation matrix. Finally, we vectorized these correlation matrices and correlated them to each other resulting in a single number (for each feature removed from the model) representing the similarity of the ``feature-removed'' models to the full model.

In order to ascertain which features were important in relating the recall models to the video model, we similarly compared to the temporal structures of video and recall models deprived of a single feature at a time (Supp. Fig.~\ref{fig:feature-importance}B). For each feature removed, we transformed each subject’s recall transcript using a model trained on the feature-deprived video text windows (of 50 scene segments), and resampled the recall timeseries to match the shape of the video model (1976 timepoints). We then represented the temporal structures of each subject’s recall model as timepoints-by -timepoints correlation matrices and computed the average correlation with the temporal structure of the video model (across subjects).
